---
title: "Milestone2"
author: "Unnati"
date: "2024-06-26"  # Replace with today's date or a specific date format
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load necessary packages
library(dplyr)
library(ggplot2)
data <-read.csv("/Users/unnatigaglani/Downloads/parkinsons_disease_data.csv")
```

# Question 1: Difference in mean MoCA scores between patients with and without Parkinson's disease
# Hypotheses:
# H0: µ1 = µ2 (Mean MoCA scores are the same for both groups)
#H1: µ1 \neq µ2 (Mean MoCA scores are different for both groups)
# Two-sample t-test

```{r Question1}
park_data <- filter(data, Diagnosis == 1) %>% select(MoCA)
no_park_data <- filter(data, Diagnosis == 0) %>% select(MoCA)
t.test(park_data$MoCA, no_park_data$MoCA, var.equal = TRUE)
```

#Interpretation
#t-value: -8.06
#Degrees of freedom (df): 2103
#p-value: 1.267e-15 (which is extremely small)
#95% confidence interval for the difference in means: -3.83 to -2.33
#Mean MoCA scores:
#Patients with Parkinson's disease (Diagnosis == 1): 13.92
#Patients without Parkinson's disease (Diagnosis == 0): 17.00
#Since the p-value is much less than 0.05, we reject the null hypothesis (H0). This indicates a significant difference in the mean MoCA scores between patients with and without Parkinson's disease. Patients without Parkinson's disease have higher mean MoCA scores.

# Question 2: Is the average UPDRS score significantly greater than 50 for patients diagnosed with Parkinson's disease?
# Hypotheses:
# H0: µ = 50 (Mean UPDRS score is equal to 50)
# H1: µ > 50 (Mean UPDRS score is greater than 50)
```{r}
updrs_data <- data %>% filter(Diagnosis == 1) %>% select(UPDRS)
t.test(updrs_data$UPDRS, mu = 50, alternative = "greater")
```
#Interpretation
# t-value: 51.559
#Degrees of freedom (df): 1303
#p-value: < 2.2e-16 (which is extremely small)
#95% confidence interval for the mean: from 116.86 to infinity
#Mean UPDRS score: 119.06
#Since the p-value is much less than 0.05, we reject the null hypothesis (H0). This indicates that the mean UPDRS score for patients with Parkinson's disease is significantly greater than 50.

# Question 3: Is there a difference in mean BMI between patients with and without a history of traumatic brain injury?
# Hypotheses:
# H0: µ1 = µ2 (Mean BMI is the same for both groups)
# H1: µ1 \neq µ2 (Mean BMI is different for both groups)
```{r}
tbi_data <- data %>% filter(TraumaticBrainInjury == 1) %>% select(BMI)
no_tbi_data <- data %>% filter(TraumaticBrainInjury == 0) %>% select(BMI)

t.test(tbi_data$BMI, no_tbi_data$BMI, var.equal = TRUE)
```
#Interpretation:
#t-value: -1.5467
#Degrees of freedom (df): 2103
#p-value: 0.1221
#95% confidence interval for the difference in means: -1.79 to 0.21
#Mean BMI values:
# Patients with a history of traumatic brain injury: 26.51
#Patients without a history of traumatic brain injury: 27.29
#Since the p-value is greater than 0.05, we fail to reject the null hypothesis (H0). This indicates that there is no significant difference in the mean BMI between patients with and without a history of traumatic brain injury.
```{r}
# Output results
print("MoCA scores comparison:")
print(t.test(park_data$MoCA, no_park_data$MoCA, var.equal = TRUE))
print("UPDRS scores comparison:")
print(t.test(updrs_data$UPDRS, mu = 50, alternative = "greater"))
print("BMI scores comparison:")
print(t.test(tbi_data$BMI, no_tbi_data$BMI, var.equal = TRUE))
```





